---
title: "LarvaTrackR workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{LarvaTrackR workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```



```{r}
library(LarvaTrackR)
library(dplyr)
```

## Step 1: Read video frames

```{r eval=FALSE}
frames <- read_video_frames(
  video_path = "~/Data/Fly_data/original.mp4",  fps = 5) # select th enumber of frame

pixel_size_mm <- 10 / 54  # check your scale here please

roi <- select_roi(
  frames,
  frame_number = 1
)     # select ROI by simply click on top-left and bottom-right

frames_roi <- crop_frames_to_roi( frames,  roi) # apply the new ROI to all frames!
```
## check the frames cut
```{r}
# test b4 going ahead
plot_frame(
  frames_roi,
  frame_number = 5
)
```


## select objects
```{r}
selected_objects <- select_objects_precise(
  frames = frames_roi,
  frame_number = 1,
  n = 4, #NUMBER OF ANIMALS
  zoom_size = 45,rotate_180 = FALSE,
  point_size = 0.5
)
```


## Visualise objects
```{r}
#check if has been assigned good
plot_selected_objects(
  frames_roi,
  selected_objects,
  frame_number = 4
)
```




## Method 1
```{r}
frame_mats <- lapply(frames_roi, get_gray_matrix)

tracks <- track_selected_points(
    frames = frames_roi,
    selected_objects = selected_objects,
    template_radius = 12,
    search_radius = 30,
    update_template = TRUE
)
plot_tracks_on_frame(tracks, frames, show_full_track = T,) # if the tracks are shitty select the second method. 
```

## method 2
```{r}
tracks <- track_selected_points_bright( frames = frames_roi,  selected_objects = selected_objects,  frame_mats = frame_mats,  search_radius = 20,  threshold_quantile = 0.96,  min_pixels = 5,  max_jump_px = 20)
plot_tracks_on_frame(tracks, frames = frames)
```

## Plot movement outputs and Calculate movement metrics
```{r}
pixel_size_mm <- 10 / 54

tracks_movement <- calculate_velocity_mm(
    tracks,
    fps = 5,
    pixel_size_mm = pixel_size_mm
)


tracks_movement_analysis <- tracks_movement %>%
  dplyr::filter(frame > 10)

tracks_movement_clean <- calculate_movement_mm_clean(
  tracks,
  fps = 5,
  pixel_size_mm = pixel_size_mm,
  min_step_px = 1.5,
  max_step_px = 34,
  remove_first_frames = 2)

plot_total_distance(tracks_movement_clean)
```


## Plot movement from the start
```{r}

pixel_size_mm <- 10 / 54

tracks_movement <- add_displacement_from_start(
  tracks_movement,
  pixel_size_mm = pixel_size_mm
)
library(ggplot2)

ggplot(tracks_movement, aes(x = time_s, y = displacement_from_start_mm,
                            color = factor(track_id))) +
  geom_line(linewidth = 1) +
  theme_classic(base_size = 14) +
  labs(
    x = "Time (s)",
    y = "Displacement from start (mm)",
    color = "Track ID"
  )

```


# tracks movement summary
```{r}
library(dplyr)

tracks_movement <- tracks_movement %>%
  arrange(track_id, frame) %>%
  group_by(track_id) %>%
  mutate(
    total_distance_mm = cumsum(ifelse(is.na(displacement_mm), 0, displacement_mm))
  ) %>%
  ungroup()

movement_summary <- tracks_movement %>%
  dplyr::group_by(track_id) %>%
  dplyr::summarise(
    total_distance_mm = max(total_distance_mm, na.rm = TRUE),
    max_displacement_from_start_mm = max(displacement_from_start_mm, na.rm = TRUE),
    mean_speed_mm_s = mean(speed_mm_s, na.rm = TRUE),
    max_speed_mm_s = max(speed_mm_s, na.rm = TRUE),
    .groups = "drop"
  )

movement_summary
```


## plot tracks simple
```{r}
a <- plot_track_simple(
  tracks = tracks,
  frames = frames_roi,
  frame_number = 1,
  flip_y_for_display = F,
  equal_aspect = FALSE
)
b <- plot_tracks_on_frame(
  frames = frames_roi,line_width = 0.4,point_size = 0.5,
  tracks = tracks,
  frame_number = 1
)

cowplot::plot_grid(a, b)
```



# plot 3D
```{r}
plot_tracks_3d(
  tracks = tracks,
  frames = frames_roi,
  fps = 5,
  first_frame = 1,
  last_frame = length(frames_roi),
  image_downsample = 4,
  flip_y_for_display = TRUE
)
p3d
```


#Companion distance analysis
```{r}
companion_distance <- calculate_companion_distance(
  tracks = tracks,
  fps = 5,
  pixel_size = pixel_size_mm
)

head(companion_distance)


plot_speed_over_time(
  tracks_movement,
  individual = TRUE,
  max_speed = 7
)

plot_companion_distance(companion_distance)
```



# plot speed over time

```{r}
tracks_movement <- calculate_movement_mm(
  tracks,
  fps = 5,
  pixel_size_mm = pixel_size_mm
)
plot_speed_over_time(tracks_movement)
```


## if you have high values. 

```{r}
tracks_movement_clean <- tracks_movement %>%
  dplyr::mutate(
    speed_mm_s_clean = dplyr::if_else(
      speed_mm_s > 10,
      NA_real_,
      speed_mm_s
    )
  )

a <- plot_speed_over_time_clean(
  tracks_movement_clean,
  individual = TRUE, max_speed = 10
)

b <- plot_speed_over_time_clean(
  tracks_movement_clean,
  individual = FALSE
)


c <- plot_tracks_on_frame(
  tracks = tracks,
  frames = frames_roi,
  frame_number = 5
)


cowplot::plot_grid(a, b, c)
```

# summary of speed 
```{r}
######## frezzing 
summary(tracks_movement_clean$speed_mm_s)

hist(
    tracks_movement_clean$speed_mm_s,
    breaks = 50,
    main = "Speed distribution",
    xlab = "Speed mm/s"
)
```

#Detect freezing-like behaviour
```{r}
#freezing = speed_mm_s < 0.8
# -----------------------------
# Add freezing status
# -----------------------------

tracks_freezing <- add_freezing_status(
  tracks_movement_clean,
  speed_threshold = 0.5,
  min_freeze_duration_s = 5,
  fps = 5,
  speed_col = "speed_mm_s_clean"
)

# -----------------------------
# Summary per animal
# -----------------------------

freezing_summary <- tracks_freezing %>%
  dplyr::group_by(track_id) %>%
  dplyr::summarise(
    total_time_s = dplyr::n() / 5,
    freezing_time_s = sum(freezing, na.rm = TRUE) / 5,
    freezing_index = freezing_time_s / total_time_s,
    freezing_percent = freezing_index * 100,
    mean_speed_mm_s = mean(speed_mm_s_clean, na.rm = TRUE),
    median_speed_mm_s = median(speed_mm_s_clean, na.rm = TRUE),
    .groups = "drop"
  )

freezing_summary
```

# freezing status plot 1
```{r}
p_freezing_index <- ggplot2::ggplot(
  freezing_summary,
  ggplot2::aes(x = "Condition", y = freezing_percent)
) +
  ggplot2::geom_boxplot(
    width = 0.35,
    outlier.shape = NA,
    linewidth = 0.6
  ) +
  ggplot2::geom_jitter(
    width = 0.08,
    size = 2.5,
    alpha = 0.8
  ) +
  ggplot2::theme_classic(base_size = 13) +
  ggplot2::labs(
    x = NULL,
    y = "Freezing index (%)",
    title = "Freezing index per animal"
  ) +
  ggplot2::theme(
    axis.text.x = ggplot2::element_blank(),
    axis.ticks.x = ggplot2::element_blank(),
    plot.title = ggplot2::element_text(face = "bold")
  )

p_freezing_index
```

# freezing status plot 2
```{r}
p_freezing_raster <- ggplot2::ggplot(
  tracks_freezing,
  ggplot2::aes(
    x = time_s,
    y = factor(track_id),
    fill = freezing
  )
) +
  ggplot2::geom_tile(height = 0.85) +
  ggplot2::scale_fill_manual(
    values = c("FALSE" = "grey85", "TRUE" = "black"),
    labels = c("Moving", "Freezing")
  ) +
  ggplot2::theme_classic(base_size = 13) +
  ggplot2::labs(
    x = "Time (s)",
    y = "Track ID",
    fill = NULL,
    title = "Freezing raster"
  ) +
  ggplot2::theme(
    plot.title = ggplot2::element_text(face = "bold"),
    legend.position = "top"
  )

p_freezing_raster
```
```{r}
p_speed_trace <- ggplot2::ggplot(
  tracks_freezing,
  ggplot2::aes(
    x = time_s,
    y = speed_mm_s_clean,
    group = track_id
  )
) +
  ggplot2::geom_line(alpha = 0.45, linewidth = 0.4) +
  ggplot2::geom_hline(
    yintercept = 0.5,
    linetype = "dashed",
    linewidth = 0.6
  ) +
  ggplot2::theme_classic(base_size = 13) +
  ggplot2::labs(
    x = "Time (s)",
    y = "Speed (mm/s)",
    title = "Speed traces with freezing threshold"
  ) +
  ggplot2::theme(
    plot.title = ggplot2::element_text(face = "bold")
  )

p_speed_trace
```

# freezing status plot 3
```{r}
freezing_bouts <- tracks_freezing %>%
  dplyr::arrange(track_id, frame) %>%
  dplyr::group_by(track_id) %>%
  dplyr::mutate(
    freezing_start = freezing == TRUE &
      dplyr::lag(freezing, default = FALSE) == FALSE,
    bout_id = cumsum(freezing_start)
  ) %>%
  dplyr::filter(freezing == TRUE) %>%
  dplyr::group_by(track_id, bout_id) %>%
  dplyr::summarise(
    bout_duration_s = dplyr::n() / 5,
    .groups = "drop"
  )

freezing_bout_summary <- freezing_bouts %>%
  dplyr::group_by(track_id) %>%
  dplyr::summarise(
    n_freezing_bouts = dplyr::n(),
    mean_bout_duration_s = mean(bout_duration_s, na.rm = TRUE),
    max_bout_duration_s = max(bout_duration_s, na.rm = TRUE),
    .groups = "drop"
  )

freezing_bout_summary

p_bout_number <- ggplot2::ggplot(
  freezing_bout_summary,
  ggplot2::aes(x = factor(track_id), y = n_freezing_bouts)
) +
  ggplot2::geom_col(width = 0.65) +
  ggplot2::theme_classic(base_size = 13) +
  ggplot2::labs(
    x = "Track ID",
    y = "Number of freezing bouts",
    title = "Freezing bouts per animal"
  ) +
  ggplot2::theme(
    plot.title = ggplot2::element_text(face = "bold")
  )

p_bout_duration <- ggplot2::ggplot(
  freezing_bouts,
  ggplot2::aes(x = factor(track_id), y = bout_duration_s)
) +
  ggplot2::geom_boxplot(outlier.shape = NA, width = 0.5) +
  ggplot2::geom_jitter(width = 0.12, size = 1.5, alpha = 0.7) +
  ggplot2::theme_classic(base_size = 13) +
  ggplot2::labs(
    x = "Track ID",
    y = "Bout duration (s)",
    title = "Freezing bout duration"
  ) +
  ggplot2::theme(
    plot.title = ggplot2::element_text(face = "bold")
  )

p_bout_number
p_bout_duration
```
```{r fig.height=12, fig.width=12}
final_freezing_plot <- cowplot::plot_grid(
  p_freezing_index,
  p_speed_trace,
  p_freezing_raster,
  p_bout_number,
  p_bout_duration,
  ncol = 2,
  labels = c("A", "B", "C", "D", "E"),
  label_size = 14
)

final_freezing_plot
```


## Notes

Interactive functions such as `select_roi()`, `calibrate_scale()`, and `select_objects()` require manual clicking on a plot. Therefore, these examples use `eval=FALSE` so the vignette can build automatically without requiring user interaction.
